---
title: "Mini Project 2"
author: "Vatsala, Emma, Margaret"
output: html_document
code_folding: hide
---

Context and Goal:
The US 2012 midterm elections took place at the same time President Barack Obama was voted into the White House for a second term. **This project seeks to understand what effect Obama’s second term as a Democrat leader would have on the 2012 House of Representatives election in New York and which play a greater role in determining the final outcome: campaign donations in support or against a candidate.** In our examination, we used data from the Federal Election Commission (FEC). More specifically, we used the *House Elections,* *Committees* and *Contributions* tables from the data set. 

Hypothesis:

1.
  a. We expect that the Democrats will win the most number of votes in the general elections as New York is Democratic. 
  b. However, the margin of votes between the Republicans and Democrats would be small because political analysts note that support for the party in power often dwindles during a president’s second term.
  c. Libertarians and conservatives will garner more votes than other third parties because mainstream media is more focused on them. 


2.
  a. More donations will be made in favor of a candidate’s campaigns than against. 
  b. The Democrats followed by the Republicans would have received the most money. 

3.
  a. Finally, there will be a positive correlation between the amount of money candidates receive in donations to use in support of their campaign and the overall votes they garner. 

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(tidyverse)
library(ggplot2)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```


```{r message = FALSE, echo = FALSE, warning = FALSE, results='hide'}
join_df <- house_elections %>%
              left_join(committees, by=c("state"= "cmte_state"))
join_df
# joining the house elections and committees dataframes to have candidates and the comittees who gave them money. We are joining by their common variable, state.
```


```{r message = FALSE, echo = FALSE, warning = FALSE,results='hide'}
fec_df<- join_df %>%
      select(state, candidate_name, party, general_votes, incumbent, cmte_name)

fec_df
#selecting only the variables that we need from the previous data frame
```


```{r message = FALSE, echo = FALSE, warning = FALSE, results='hide'}
contributions_df <- contributions %>%
  select(name,transaction_amt,transaction_type, state)%>%
  filter(transaction_amt >0)

contributions_df
# Selecting the useful variables from the contributions data frame and then filtering to only get contributions that werent returned. 
```


```{r message = FALSE, echo = FALSE, warning = FALSE, results='hide'}
contributions_fec <- fec_df %>%
    left_join(contributions_df, by = c("cmte_name" = "name"))%>%
    filter(!is.na(transaction_amt)) %>%
  group_by(state.x) %>%
  arrange(state.x)

contributions_fec
# joining the contributions data frame and the fec data frame by name. We are filtering to remove the NA values and grouping by and arranging by state.
```



```{r message = FALSE, echo = FALSE, warning = FALSE, results='hide'}
sum_money<- contributions_fec %>%
  group_by(cmte_name) %>%
  summarize( sum_money = sum(transaction_amt))

sum_money
# finding the sum of each committees donations
```



```{r message = FALSE, echo = FALSE, warning = FALSE}
general_votes_df <- contributions_fec %>%
  left_join(sum_money, by = "cmte_name") %>%
  filter(transaction_type==c("24E", "24A"),state.x == "NY") %>%
  select(state.x, party, general_votes, incumbent, candidate_name, cmte_name, transaction_amt, transaction_type, sum_money)

general_votes_df
#Joining the data sets so we can have each candidate, their status of imcumbancy, the total amount of money they received from each committee, and the total amount of general votes they received. We are filtering so we have both how much money was used against the candidate and how much mone was used in their favor. We can use this to see each party and how many votes they received. 
```


```{r message = FALSE, echo = FALSE, warning = FALSE} 
donations_df <- contributions_fec %>%
  left_join(sum_money, by = "cmte_name")%>%
  filter(transaction_type==c("24E", "24A"), state.x == "NY")%>%
  select(state.x, candidate_name, party, general_votes, incumbent, transaction_type, sum_money)%>%
  group_by(transaction_type, party)%>%
  summarise(new_sum=sum(sum_money))

donations_df
# Doing the same thing as the data frame above but now groupnig by the transaction type. We want to see how much money was used both in favor of a party and against a party. 
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
general_votes_plot <- ggplot(general_votes_df, aes(x = party, y = general_votes)) + 
  geom_col(position = position_dodge(), width=0.5,fill="#A4A4A4")+
   scale_x_discrete(label = c("Conservative","Democrat","Green", "Libertarian","Republican","Socialist", "Write In"))+
  labs(title = "Total General Votes Each Party Received", x = "Party", y = "General Votes")+
  theme_minimal()
  

general_votes_plot
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
donations_plot <- ggplot(donations_df, aes(x = party, y = new_sum, fill = transaction_type))+
  geom_col(position = position_dodge())+
     scale_x_discrete(label = c("Conservative","Democrat","Green", "Libertarian","Republican","Socialist", "Write In"))+
   scale_fill_manual(name="Transaction Type",values = c("#CC6666", "#9999CC"),labels=c("Donations made AGAINST a candidate", "Donations made FOR a candidate"))+
  theme_minimal()+
  theme(legend.position = "top")+
  labs(title = "Amount of Money used for and Against each Party", y = "Amount of Money", x = "Party")

donations_plot
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
votes_function<- function(state_arg = "NY", money_arg1 = "24A",money_arg2 = "24E"){
  general_votes_df <- contributions_fec %>%
  left_join(sum_money, by = "cmte_name") %>%
  filter(transaction_type== c(money_arg1, money_arg2), state.x == state_arg) %>%
  select(state.x, party, general_votes, incumbent, candidate_name, cmte_name, transaction_amt, transaction_type, sum_money)
  general_votes_df
}
# function to iterate over any state
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
donations_function <- function(state_arg = "NY",money_arg1 = "24A",money_arg2 = "24E"){
  donations_df <- contributions_fec %>%
  left_join(sum_money, by = "cmte_name")%>%
  filter(transaction_type==c(money_arg1, money_arg2), state.x == state_arg)%>%
  select(state.x, candidate_name, party, general_votes, incumbent, transaction_type, sum_money)%>%
  group_by(transaction_type, party)%>%
  summarise(new_sum=sum(sum_money))

donations_df
}
# function to iterate over any state or donation/transaction type
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(devtools)
devtools::install_github("benmarwick/wordcountaddin", type = "source", dependencies = TRUE)
```



Analysis: 

1. 
  a. From the first graph we can see that the Democrats received the majority of votes in New York, followed by the Republicans. New York is a historically Democratic state and is expected to remain that way. Moreover, in 2012 almost half the registered voters were Democrats (Wheaton).
  b. We correctly hypothesized that the margin of votes between the Republicans and Democrats would not be high. This can be explained by voter turnouts as only the most passionate voters participate in midterm elections. Of all voters, Republicans who feel they are a minority in New York are more likely to vote in the midterm elections than Democrats (Pew Research Center). Furthermore, fewer registered Democrats may have voted because of a sense of confidence in the party’s success created by Obama’s presidential victory. Therefore, Obama’s presidential victory may have caused a more Republican outcome in the midterms. 
  c. It is interesting to note that neither the Libertarians nor the Conservatives formed the largest third party, the Green party did. Libertarians received the least number of votes, despite more media focus on them including that of the write-ins who were not even on the ballot. Also, the Conservative party received the same number of votes as the Socialist party. This is surprising because it is generally thought that the Socialists only win votes in times of socio-economic crisis, which was not the case in 2012 (Amadeo). 


2.
  a. As was hypothesized, more money was donated in support of candidates than against; it makes more sense for individuals to fund those they support rather than donate to oppose another. What was interesting was that the same amount of money was donated against write-ins, Conservatives and Libertarians suggesting that all three parties are considered equally threatening to donors of New York. 
  b. The Democrats and Republicans received the greatest number of financial campaign contributions. This was as expected because they are the largest parties in the state. 


3.
  a. There is a positive relationship between the amount of donations received and the number of votes garnered for the Democrat, Republican, and Green parties. However, this does not apply for the other third parties. Conservatives, Libertarians and Write-ins received more donations than the Socialists, but the Socialists garnered more votes than both the Libertarians and Write-ins and about the same as the Conservatives. From this we see that there are other variables outside of campaign financing and historical trends that affect the number of votes candidates and parties receive. 
  
Sources:

Wheaton, Sarah. “New York: Not as Blue as it Could Be”, The New York Times,   27  July 2012, https://fivethirtyeight.blogs.nytimes.com/2012/07/27/new-york-      not-as-blue-as-it-could-be/

“Who Votes, Who Doesn’t, and Why:Regular Voters, Intermittent Voters, and       Those Who Don't”, Pew Research Center, 18 October 2006,                       https://www.people-press.org/2006/10/18/who-votes-who-doesnt-and-why/

Amadeo, Kimberly. “US Economy 2012: Summary and Critical Events”, The          Balance, 25 January 2019,https://www.thebalance.com/u-s-economy-2012-330574

*^[https://github.com/v-ramanan/sds192-mp2.git]



