---
title: "Mini Project 2"
author: "Vatsala, Emma, Margaret"
output: html_document
code_folding: hide
---

Context and Goal:
The US 2012 midterm elections took place at the same time President Barack Obama (Democrat) was voted into the White House for a second term. Through our project, we sought to understand what effect Obama’s second term as a Democrat leader would have on the 2012 House of Representatives election in the large and historically Democratic state of New York and which play a greater role in determining the final outcome: campaign donations in support or against a candidate. In our examination, we used data from the Federal Election Commission (FEC). More specifically, we used the House Elections, Committees and Contributions tables from the data set. 

Hypothesis:
1. 
  + 1a It is expected that the Democrats will win the most number of votes in the general elections as New York is a historically blue state.
  + 1b However, the margin of votes between the Republicans and Democrats would be small because political analysts note that support for the party in power often dwindles during a president’s second term (citation needed).
  + 1c Libertarians and conservatives will garner more votes than other third parties because mainstream political media is more focused on them. 


2. 
  + 2a More donations will be made in favor of a candidate’s campaigns than against. 
  + 2b The Democrats followed by the Republicans would have received the most money. 


3.
  + 3a Finally, there will be a positive correlation between the amount of money candidates receive in donations to use in support of their campaign and the overall votes they garner. 

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(tidyverse)
library(ggplot2)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```


```{r message = FALSE, echo = FALSE, warning = FALSE, results='hide'}
join_df <- house_elections %>%
              left_join(committees, by=c("state"= "cmte_state"))
join_df
# joining the house elections and committees dataframes to have candidates and the comittees who gave them money. We are joining by their common variable, state.
```


```{r message = FALSE, echo = FALSE, warning = FALSE,results='hide'}
fec_df<- join_df %>%
      select(state, candidate_name, party, general_votes, incumbent, cmte_name)

fec_df
#selecting only the variables that we need from the previous data frame
```


```{r message = FALSE, echo = FALSE, warning = FALSE, results='hide'}
contributions_df <- contributions %>%
  select(name,transaction_amt,transaction_type, state)%>%
  filter(transaction_amt >0)

contributions_df
# Selecting the useful variables from the contributions data frame and then filtering to only get contributions that werent returned. 
```


```{r message = FALSE, echo = FALSE, warning = FALSE, results='hide'}
contributions_fec <- fec_df %>%
    left_join(contributions_df, by = c("cmte_name" = "name"))%>%
    filter(!is.na(transaction_amt)) %>%
  group_by(state.x) %>%
  arrange(state.x)

contributions_fec
# joining the contributions data frame and the fec data frame by name. We are filtering to remove the NA values and grouping by and arranging by state.
```



```{r message = FALSE, echo = FALSE, warning = FALSE, results='hide'}
sum_money<- contributions_fec %>%
  group_by(cmte_name) %>%
  summarize( sum_money = sum(transaction_amt))

sum_money
# finding the sum of each committees donations
```



```{r message = FALSE, echo = FALSE, warning = FALSE}
general_votes_df <- contributions_fec %>%
  left_join(sum_money, by = "cmte_name") %>%
  filter(transaction_type==c("24E", "24A"),state.x == "NY") %>%
  select(state.x, party, general_votes, incumbent, candidate_name, cmte_name, transaction_amt, transaction_type, sum_money)

general_votes_df
#Joining the data sets so we can have each candidate, their status of imcumbancy, the total amount of money they received from each committee, and the total amount of general votes they received. We are filtering so we have both how much money was used against the candidate and how much mone was used in their favor. We can use this to see each party and how many votes they received. 
```


```{r message = FALSE, echo = FALSE, warning = FALSE} 
donations_df <- contributions_fec %>%
  left_join(sum_money, by = "cmte_name")%>%
  filter(transaction_type==c("24E", "24A"), state.x == "NY")%>%
  select(state.x, candidate_name, party, general_votes, incumbent, transaction_type, sum_money)%>%
  group_by(transaction_type, party)%>%
  summarise(new_sum=sum(sum_money))

donations_df
# Doing the same thing as the data frame above but now groupnig by the transaction type. We want to see how much money was used both in favor of a party and against a party. 
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
general_votes_plot <- ggplot(general_votes_df, aes(x = party, y = general_votes)) + 
  geom_col(position = position_dodge(), width=0.5,fill="#A4A4A4")+
   scale_x_discrete(label = c("Conservative","Democrat","Green", "Libertarian","Republican","Socialist", "Write In"))+
  labs(title = "Total General Votes Each Party Received", x = "Party", y = "General Votes")+
  theme_minimal()
  

general_votes_plot
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
donations_plot <- ggplot(donations_df, aes(x = party, y = new_sum, fill = transaction_type))+
  geom_col(position = position_dodge())+
     scale_x_discrete(label = c("Conservative","Democrat","Green", "Libertarian","Republican","Socialist", "Write In"))+
   scale_fill_manual(name="Transaction Type",values = c("#CC6666", "#9999CC"),labels=c("Donations made AGAINST a candidate", "Donations made FOR a candidate"))+
  theme_minimal()+
  theme(legend.position = "top")+
  labs(title = "Amount of Money used for and Against each Party", y = "Amount of Money", x = "Party")

donations_plot
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
votes_function<- function(state_arg = "NY", money_arg1 = "24A",money_arg2 = "24E"){
  general_votes_df <- contributions_fec %>%
  left_join(sum_money, by = "cmte_name") %>%
  filter(transaction_type== c(money_arg1, money_arg2), state.x == state_arg) %>%
  select(state.x, party, general_votes, incumbent, candidate_name, cmte_name, transaction_amt, transaction_type, sum_money)
  general_votes_df
}
# function to iterate over any state
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
donations_function <- function(state_arg = "NY",money_arg1 = "24A",money_arg2 = "24E"){
  donations_df <- contributions_fec %>%
  left_join(sum_money, by = "cmte_name")%>%
  filter(transaction_type==c(money_arg1, money_arg2), state.x == state_arg)%>%
  select(state.x, candidate_name, party, general_votes, incumbent, transaction_type, sum_money)%>%
  group_by(transaction_type, party)%>%
  summarise(new_sum=sum(sum_money))

donations_df
}
# function to iterate over any state or donation/transaction type
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(devtools)
devtools::install_github("benmarwick/wordcountaddin", type = "source", dependencies = TRUE)
```



Analysis: 

1. 
  +a From the first graph we can see that the Democrats received the majority of votes in New York, followed by the Republicans. This was as hypothesised; New York is a historically Democratic state and, unless a major social, political or economic event occurs, it is expected to remain that way. Moreover, in 2012 almost half the registered voters were democrats(Wheaton).
  +b It was also correctly hypothesised that the margin of votes between the Republicans and Democrats would not be that extensive. One can understand this occurrence by considering voter turnout. It must be noted that only the most passionate voters participate in midterm elections. In 2006, the most passionate and regular voter comprised 35% of voters (Pew Research Center). Fewer registered Democrats may have voted in 2012 because of a sense of confidence in the party’s success created by Obama’s presidential victory. However, history is also at play in this scenario. Since New York is a historically blue state, Republican voters who feel they are a minority in New York are more likely to vote in the midterm elections than more moderate Democrats (Pew Research Center).Therefore, Obama’s presidential victory may have caused a more Republican outcome of the midterms. 
  +c It is interesting to note that that neither the Libertarians nor the Conservatives formed the largest third party; the Green party garnered the third most votes. Libertarians, in fact, received the least number of votes, despite more media focus on them, among all the  third parties, including that of the write-ins who were not even on the ballot. We also see that the Conservative party received the same number of votes as the Socialist party. This is surprising because it is generally thought that the Socialist party only wins votes in times of major socio-economic crisis, which was not the case in 2012 (Amadeo). 


2.
  +a As was hypothesised, more money was donated in support of candidates than against as it would make more sense for individuals to fund those they support rather than donate to oppose another. What was interesting was that the same amount of money was donated against write-in candidates, who do not even appear on the ballot, as was against Conservatives and Libertarians suggesting that all three parties are considered equally threatening to donors of New York. 
  +b The Democrats and Republicans received the greatest number of financial campaign contributions, as compared to other third parties. From this, more of the money was spent in support of candidates, rather than against them. This was as expected because, being the largest parties, the Democrats and Republicans would naturally receive the most money.


3.
  +a When considering both the graphs together, one can see that there is a positive relationship between the amount of donations received and the number of votes garnered for the Democrat, Republican, and Green parties. However, this does not apply for the other third parties. Conservatives, Libertarians and Write-ins received more donations than the Socialists, but the Socialists garnered more votes than both the Libertarians and Write-ins and about the same as the Conservatives. From this we see that there are other variables outside of campaign financing and historical trends that affect the number of votes candidates and parties receive. 


*^[https://github.com/v-ramanan/sds192-mp2.git]
*^[Wheaton, Sarah. “New York: Not as Blue as it Could Be”, The New York Times, 27 July 2012, https://fivethirtyeight.blogs.nytimes.com/2012/07/27/new-york-not-as-blue-as-it-could-be/
]
*^[“Who Votes, Who Doesn’t, and Why:Regular Voters, Intermittent Voters, and Those Who Don't”, Pew Research Center, 18 October 2006, https://www.people-press.org/2006/10/18/who-votes-who-doesnt-and-why/]
*^[Amadeo, Kimberly. “US Economy 2012: Summary and Critical Events”, The Balance, 25 January 2019,https://www.thebalance.com/u-s-economy-2012-330574
]


